Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-01-14 Grassmannian_cells.tex
Section 3.2 edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}

@@

In this section, we fix compatible bases for each $H_m=\Hom(P_{m-1},P_m)$ and use these to lift the (truncated) preprojectives to the universal cover of $K(n)$.
This lifting will rigidify the situation and allow more precise control over these representations and their subrepresentation structure.
The main advantage of this lifting is that those truncated preprojectives which can be lifted are exceptional representations on the universal covering.

Fix a natural number $n\geq 3$ and consider the following lifts $\tilde P_m$  of $P_m$ to the universal cover $\widetilde{K(n)}$.
For $m=1$, we assume that $\dim(\tilde P_1)_{1,1}=1$ and $\dim (\tilde P_1)_{i,w}=0$ if $(i,w)\neq (1,1)$.
Moreover, we assume that $\dim (\tilde P_2)_{2,1}=\dim (\tilde P_2)_{1,w}=1$ for $w=\alpha_1,\ldots,\alpha_n$ and $\dim (\tilde P_2)_{i,w}=0$ if $(i,w)\notin\{(2,1),(1,\alpha_1),\ldots,(1,\alpha_n)\}$.
Recall that applying the BGP-reflection functor $\Sigma_i$ on $K(n)$ corresponds to applying $\tilde\Sigma_i:=\prod_{w\in W_Q}\Sigma_{(i,w)}$ on $\widetilde{K(n)}$.
Moreover, sinks become sources and vice versa.
For $m\geq 2$, we build lifts $\tilde P_m$ recursively by applying reflection functors or as Auslander-Reiten translates, i.e. we have
\begin{align}
  \label{recursive}
  \tilde P_m=(\tilde\Sigma_2 \tilde P_{m-1})^{\sigma},\quad\tilde P_{m+1}=\tilde\Sigma_{1}\tilde\Sigma_2\tilde P_{m-1}=\tau^{-1} \tilde P_{m-1}.
\end{align}

It is rather important that $\tilde P_{2l}$ and $\tilde P_{2l-1}$ live on two different components of the universal covering quiver.

\begin{remark}
  \mbox{}
  \begin{enumerate}
    \item Note that the group $W_Q$ acts naturally on $\widetilde{K(n)}_0$ via translation, i.e. $w.(i,w')=(i,ww')$, and this induces an action of $W_Q$ on $\rep\widetilde{K(n)}$. 
      %Up to this action a lift $\tilde P_m$ of $P_m$ is unique (on isomorphism classes).
    \item  Every representation $\tilde P_m$ has $n$ subrepresentation $\tilde P_{m-1,1},\ldots,\tilde P_{m-1,n}$. Each of them corresponds to one of the $n$ different arrows of $K(n)$. 	
		We have 
      $$\tilde P_{2l-1,i}:=\alpha_i.\tilde P_{2l-1},\quad \tilde P_{2l,i}:=\alpha_i^{-1}.\tilde P_{2l}$$
      for all $l\geq 1$ and for $i=1,\ldots,n$. This can be checked straightforwardly for $l=1$ and follows in general when applying the identities (\ref{recursive}).

    \item Note that, the dimension vector of $\tilde P_m$ are symmetric. They all have the same middle vertex $(1,1)$ if $m$ is odd and $(2,1)$ if $m$ is even.
\end{enumerate}
\end{remark}
Let
$$I^n_l:=\{(j_1,\ldots,j_{l})\mid j_i\in\{1,\ldots,n\},\,j_i\neq j_k\text{ if }i\neq k\}.$$
As we fixed the number of arrows, we drop the $n$ in what follows.

@@

\begin{remark}
Actually, these examples already show how the (truncated) preprojectives are built, also using the identities (\ref{recursive}). We should point out that the case $m=3$ is rather special because it is the only case for which the unique morphism $\tilde P_{2,j_l}\to\tau\tilde P_3^{\uj}$ is not surjective. Note that if we factor out the image of this morphism, the remaining representation is a direct sum of simple injective representations which disappear after reflecting at the source.
\end{remark}

@@

    %\item Finally, we have $\Ext(\tilde P_{m+1}^{(k-1)},V)=0$ for every subrepresentation $V$ of $\tilde P_{m-1}^{n-1}\oplus \tilde P_{m-2}^{n-k}$.
  \end{enumerate}%\footnote{introduce/consider/check indices, we have to mod out the corresponding reps in the covering} 
\end{lemma}
\begin{proof}
Using the identities (\ref{recursive}), it suffices to prove all statements for $m=4$ and arbitrary $1\leq l\leq n-1$. Without loss of generality, we choose $(j_1,\ldots,j_l)=(1,\ldots,l)$ and define $\tilde P_m^{(0)}=\tilde P_m$ respectively. We should point out that it is very important for this proof that we consider fixed lifts of all representations in the universal covering as defined in the beginning of this section. Note that, starting with these cases, we only reflect at sources and no representation appearing is isomorphic to the simple injective representation $S_2$.

In Example \ref{smallm} it was shown that $\tau \tilde P_4^{(1,\ldots,l)}=\tilde P_2^{(1,\ldots,l)}$. For $m=5$,  we consider the short exact sequence
\[\ses{\bigoplus_{i=1}^l\tilde P_{2,i}}{\tilde P_3}{\tilde P_3^{(1,\ldots,l)}}\]
and apply the reflection functor $\tilde\Sigma_1\tilde\Sigma_2$ which is the same as applying $\tau^{-1}$. As we have $\tau^{-1} \tilde P_{2,i}=\tilde P_{4,i}$ and $\tau^{-1}\tilde P_3=\tilde P_5$, it follows that $\tau^{-1}\tilde P_3^{(1,\ldots,l)}=\tilde P_5^{(1,\ldots,l)}$ which is statement for $m=5$. The general statement follows from the identities (\ref{recursive}). Note that this also shows that
\[(\tilde \Sigma_2\tilde P_{m}^{(1,\ldots,l)})^\sigma=\tau^{-1}\tilde P^{(1,\ldots,l)}_{m-1}.\]

Considering the quiver $\tilde P_{3,l}$ from Example \ref{smallm}, it is clear that we obtain an exact sequence
$$\ses{P(4,(1,\ldots,l))}{P_{3,l}}{\tilde P_2^{(1,\ldots,l)}}.$$
Actually, the first summand of $P(4,(1,\ldots,l))$ corresponds to the quivers $\tilde A_i$ at the bottom and the second summand to the $l-1$ sources in the second row lying on the left. Together with the first statement and the identities (\ref{recursive}), this shows the second statement.

@@

We have $\Hom(\tilde P_{m}^{\uj},\tilde P_{m-1,j_l})=0$ by Lemma \ref{le:basic homological properties}. As $\Ext(\tilde P_{4}^{\uj},\tilde P_{3,j_l})=\CC$ by a straightforward check, applying (\ref{recursive}), it follows that $\Ext(\tilde P_{m}^{\uj},\tilde P_{m-1,j_l})=\CC$. This yields the third and forth identity.

@@

%Thus we see that $\tilde P_{m}^{\ui}\cong \tilde P_{m-1}^{(l)}$ for some $l\in\{1,\ldots,n\}$. But now it is straightforward to check that $l=n$. \footnote{more precise?} 
For $m=3$, it is straightforward that the injection of the lower row factors over $\tilde P_{1,(j^c,j^c)}$ which means that we get the same commutative diagram when replacing $\tilde P_{1}$ by $\tilde P_{1,(j^c,j^c)}$. Thus we have $\tilde P_{3}^{\uj}\cong \tilde P_{2,j^c}^{\uj^c}/\tilde P_{1,(j^c,j^c)}=\tilde P_{2,j^c}^{(j^c)}$. Applying the identities (\ref{recursive}) the claim follows for $m\geq 3$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}
\label{sec:covering}

@@

Fix a natural number $n\geq 3$ and write $W_n:=W_{K(n)}$ for the free group generated by the arrows $\alpha_i$, $1\le i\le n$, of $K(n)$.
In this section, we fix compatible bases for each $H_m:=\Hom(P_m,P_{m+1})$ and use these to lift the (truncated) preprojectives to the universal cover $\widetilde{K(n)}$ of the quiver $K(n)$.
This lifting will rigidify the situation, allowing more precise control over these representations and their subrepresentation structure.
The main advantage of this lifting is that those truncated preprojectives which can be lifted are exceptional representations on the universal covering quiver $\widetilde{K(n)}$.

We will mainly be interested in particular lifts $\tilde P_m$ of the preprojective representations $P_m$ of $K(n)$ to the universal cover $\widetilde{K(n)}$.
First, recall that applying the BGP-reflection functor $\Sigma_i$ on $K(n)$ corresponds to applying the iterated reflection $\tilde\Sigma_i:=\prod_{w\in W_Q}\Sigma_{(i,w)}$ on $\widetilde{K(n)}$.
Moreover, under this operation all sinks become sources and vice versa.
The preprojective lifts we use are defined recursively as follows.
\begin{itemize}
  \item For $m=1$, we consider the lift $\tilde P_1$ satisfying $\dim(\tilde P_1)_{(1,1)}=1$ and $\dim (\tilde P_1)_{(i,w)}=0$ for $(i,w)\neq (1,1)$.
  \item For $m=2$, we consider the lift $\tilde P_2$ satisfying $\dim (\tilde P_2)_{(2,1)}=\dim (\tilde P_2)_{(1,\alpha_i)}=1$ for $1\le i\le n$ and $\dim(\tilde P_2)_{(i,w)}=0$ for $(i,w)\notin\{(2,1),(1,\alpha_1),\ldots,(1,\alpha_n)\}$.
  \item For $m\geq 3$, we build the lifts $\tilde P_m$ recursively by applying reflection functors or as Auslander-Reiten translates.
    More precisely, we set
    \begin{align}
      \label{eq:recursive covers}
      \tilde P_m:=(\tilde\Sigma_2 \tilde P_{m-1})^{\sigma}\qquad\text{or}\qquad\tilde P_{m+1}:=\tilde\Sigma_{1}\tilde\Sigma_2\tilde P_{m-1}=\tau^{-1} \tilde P_{m-1}.
    \end{align}
\end{itemize}

It will be rather important that the lifts $\tilde P_{2l}$ and $\tilde P_{2l-1}$ live on two different components of the universal covering quiver.
Indeed, recall that the group $W_n$ naturally acts on $\widetilde{K(n)}_0$ via translation, i.e. $w.(i,w')=(i,ww')$, and this induces an action of $W_n$ on $\rep\widetilde{K(n)}$.
Following the notation of Section~\ref{sec:covering}, we write $\tilde P_{m,w}$ for the representation of $\widetilde{K(n)}$ obtained by translating the lifted preprojective representation $\tilde P_m$ by the action of $w\in W_n$.
To simplify the notation, we abbreviate $P_{2l-1,i}:=P_{2l-1,\alpha_i}$ and $P_{2l,i}:=P_{2l,\alpha_i^{-1}}$.
\begin{lemma}
  For each $m\ge2$, the representation $\tilde P_m$ has precisely $n$ subrepresentations covering $P_{m-1}$.
  These are the representations $P_{m-1,i}$ corresponding the $n$ different arrows of $K(n)$. 
\end{lemma}
\begin{proof}
  This is clear for $m=2$ and follows in general by applying the identities \eqref{eq:recursive covers}.
\end{proof}
\begin{remark}
  Note that the dimension vectors of the lifted preprojectives $\tilde P_m$ are symmetric.
  In particular, they all have the same middle vertex $(1,1)$ if $m$ is odd and middle vertex $(2,1)$ if $m$ is even.
\end{remark}
Let
$$I^n_l:=\{(j_1,\ldots,j_{l})\mid j_i\in\{1,\ldots,n\},\,j_i\neq j_k\text{ if }i\neq k\}.$$
As we fixed the number of arrows, we drop the $n$ in what follows.

@@

\begin{remark}
Actually, these examples already show how the (truncated) preprojectives are built, also using the identities \eqref{eq:recursive covers}. We should point out that the case $m=3$ is rather special because it is the only case for which the unique morphism $\tilde P_{2,j_l}\to\tau\tilde P_3^{\uj}$ is not surjective. Note that if we factor out the image of this morphism, the remaining representation is a direct sum of simple injective representations which disappear after reflecting at the source.
\end{remark}

@@

    %\item Finally, we have $\Ext(\tilde P_{m+1}^{(k-1)},V)=0$ for every subrepresentation $V$ of $\tilde P_{m-1}^{n-1}\oplus \tilde P_{m-2}^{n-k}$.
  \end{enumerate}%\footnote{introduce/consider/check indices, we have to mod out the corresponding reps in the covering} 
\end{lemma}
\begin{proof}
Using the identities \ref{eq:recursive covers}, it suffices to prove all statements for $m=4$ and arbitrary $1\leq l\leq n-1$. Without loss of generality, we choose $(j_1,\ldots,j_l)=(1,\ldots,l)$ and define $\tilde P_m^{(0)}=\tilde P_m$ respectively. We should point out that it is very important for this proof that we consider fixed lifts of all representations in the universal covering as defined in the beginning of this section. Note that, starting with these cases, we only reflect at sources and no representation appearing is isomorphic to the simple injective representation $S_2$.

In Example \ref{smallm} it was shown that $\tau \tilde P_4^{(1,\ldots,l)}=\tilde P_2^{(1,\ldots,l)}$. For $m=5$,  we consider the short exact sequence
\[\ses{\bigoplus_{i=1}^l\tilde P_{2,i}}{\tilde P_3}{\tilde P_3^{(1,\ldots,l)}}\]
and apply the reflection functor $\tilde\Sigma_1\tilde\Sigma_2$ which is the same as applying $\tau^{-1}$. As we have $\tau^{-1} \tilde P_{2,i}=\tilde P_{4,i}$ and $\tau^{-1}\tilde P_3=\tilde P_5$, it follows that $\tau^{-1}\tilde P_3^{(1,\ldots,l)}=\tilde P_5^{(1,\ldots,l)}$ which is statement for $m=5$. The general statement follows from the identities \ref{eq:recursive covers}. Note that this also shows that
\[(\tilde \Sigma_2\tilde P_{m}^{(1,\ldots,l)})^\sigma=\tau^{-1}\tilde P^{(1,\ldots,l)}_{m-1}.\]

Considering the quiver $\tilde P_{3,l}$ from Example \ref{smallm}, it is clear that we obtain an exact sequence
$$\ses{P(4,(1,\ldots,l))}{P_{3,l}}{\tilde P_2^{(1,\ldots,l)}}.$$
Actually, the first summand of $P(4,(1,\ldots,l))$ corresponds to the quivers $\tilde A_i$ at the bottom and the second summand to the $l-1$ sources in the second row lying on the left. Together with the first statement and the identities \eqref{eq:recursive covers}, this shows the second statement.

@@

We have $\Hom(\tilde P_{m}^{\uj},\tilde P_{m-1,j_l})=0$ by Lemma \ref{le:basic homological properties}. As $\Ext(\tilde P_{4}^{\uj},\tilde P_{3,j_l})=\CC$ by a straightforward check, applying \ref{eq:recursive covers}, it follows that $\Ext(\tilde P_{m}^{\uj},\tilde P_{m-1,j_l})=\CC$. This yields the third and forth identity.

@@

%Thus we see that $\tilde P_{m}^{\ui}\cong \tilde P_{m-1}^{(l)}$ for some $l\in\{1,\ldots,n\}$. But now it is straightforward to check that $l=n$. \footnote{more precise?} 
For $m=3$, it is straightforward that the injection of the lower row factors over $\tilde P_{1,(j^c,j^c)}$ which means that we get the same commutative diagram when replacing $\tilde P_{1}$ by $\tilde P_{1,(j^c,j^c)}$. Thus we have $\tilde P_{3}^{\uj}\cong \tilde P_{2,j^c}^{\uj^c}/\tilde P_{1,(j^c,j^c)}=\tilde P_{2,j^c}^{(j^c)}$. Applying the identities \eqref{eq:recursive covers} the claim follows for $m\geq 3$.
